Tree size, exposure, and hydraulic traits interactively shape drought response in a temperate broadleaf forest

### 2019-06-29 manuscript/03-methods.Rmd
Update 03-methods.Rmd
Update 03-methods.Rmd
removed  canopy-subcanopy table, description of quilts analysis.

## Data Collection
We analyzed tree-ring data from the twelve species contributing most to woody aboveground net primary productivity (ANPP_stem) (Table 1), which together comprise 97% of whole-ecosystem ANPP_stem. Cores were collected in 2010-2011 or 2016-2017 from a breast height of 1.3m using a 5mm increment borer. In 2010-2011, cores were collected from randomly selected live trees were selected at random from each species in 2010-2011, with at least 30 of those trees having a diameter at breast height (DBH) of at least 10cm [@bourg_initial_2013]. In 2016-2017, cores were collected from all dead trees found in the annual mortality census [@gonzalezakre_patterns_2016]. Cores were sanded, measured, and cross-dated using standard procedures, as detailed in [@helcoski_growing_2019].

|**Species**|**n_cores.canopy**|**n_cores.subcanopy**|**included**
|----------|-------------|---------------|-----------|
*Carya cordiformis*|2|11|yes
*Carya glabra*|9|22|yes
*Carya ovalis*|9|14|yes
*Carya tomentosa*|0|13|sub-canopy only
*Fagus grandifolia*|7|73|yes
*Fraxinus americana*|20|42|yes
*Fraxinus nigra*|1|11| sub-canopy only
*Juglans nigra*|22|9|yes
*Liriodendron tulipifera*|38|60|yes
*Quercus alba*|38|23|yes
*Quercus prinus*|27|32|yes
*Quercus rubra*|43|26|yes
*Quercus velutina*|54|23|yes

``` {r sample, echo=FALSE}
#numbers for the table above come from this table
# library(RCurl)
# library(dplyr)
# species <- read.csv(text=getURL("https://raw.githubusercontent.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/master/data/core_list_for_neil.csv?token=AJNRBELCYZY576ARLE7BXNK5BEF3O"))
# 
# test <- species %>%
#   group_by(sp, crown.position) %>%
#   summarize(count=n())

Using CRU data **with Valentine's quilts for each grouping**, it was found that all climate variables tested were significant by tree position when using a linear mixed model (**this was done early-on, included here because unsure if will keep for full analysis**).

>>>>

## Data Collection
We analyzed tree-ring data from the twelve species contributing most to woody aboveground net primary productivity (ANPP_stem) (Table 1), which together comprise 97% of whole-ecosystem ANPP_stem. Cores were collected in 2010-2011 or 2016-2017 from a breast height of 1.3m using a 5mm increment borer. In 2010-2011, cores were collected from randomly selected live trees were selected at random from each species in 2010-2011, with at least 30 of those trees having a diameter at breast height (DBH) of at least 10cm [@bourg_initial_2013]. In 2016-2017, cores were collected from all dead trees found in the annual mortality census [@gonzalezakre_patterns_2016]. Cores were sanded, measured, and cross-dated using standard procedures, as detailed in [@helcoski_growing_2019]. The resulting chronologies have been published in association with [@helcoski_growing_2019]: (GitHub URL), (ITRDB).

*Table 1. Species analyzed here, listed in descending order of ANPP_stem. n cores and DBH range represented, and species traits* [this should combine the two remaining tables in this section]
